---
title: "multicollinearity test"
output: html_document
---

## Load data && Preprocess
```{r setup, include=FALSE}
data=read.csv("/Users/violaliu/Downloads/BodyFat.csv")
attach(data)
print(summary(data))
```

### From the summary, we can speculate that there may be several outliers due to inaccuracy of measurement. We took a closer look at the specific samples including outliers. (Then we checked BMI recorded with weight and height recorded among those three people to see if there is any obvious error.-No obvious error.) We delete the sample with bodyfat=0.0 

```{r}
data[data$AGE==81|data$BODYFAT==45.10|data$BODYFAT==45.10|data$BODYFAT==0.00|data$WEIGHT==363.15, ]

data_removed=data[data$BODYFAT!=0.0,]
```

## Correlation between independent variables

### VIF
#### check sets of variables, to see if they are independendent(can pass VIF test)
##### vif of remained feature density, abdomen, adiposity, age, chest
```{r }
model_vif_2<-lm(BODYFAT~ AGE+CHEST+ABDOMEN+DENSITY +ADIPOSITY,data=data_removed)

ols_vif_tol(model_vif_2)

```
#### delete abdomen
