---
title: "stat628-proj2-regression"
output:
  pdf_document: default
  html_document: default
---
## Correlation between independent variables

### VIF
#### check sets of variables, to see if they are independendent(can pass VIF test)
##### all variables
```{r }
library(olsrr)
model_vif<-lm(BODYFAT~ AGE+WEIGHT + HEIGHT+ADIPOSITY+NECK+CHEST+ABDOMEN+HIP+THIGH+KNEE+ANKLE+BICEPS+FOREARM+WRIST,data=data_removed)

ols_vif_tol(model_vif)

```
##### vif of remained feature density, abdomen, adiposity, age, chest
#### abdomen delete
```{r }
model_vif_2<-lm(BODYFAT~ AGE+CHEST+ABDOMEN+DENSITY +ADIPOSITY,data=data_removed)

ols_vif_tol(model_vif_2)

```

##### keep weight, remove other variables, of which VIF above 10
```{r }
model_vif_2<-lm(BODYFAT~ AGE+WEIGHT + HEIGHT+NECK+KNEE+ANKLE+BICEPS+FOREARM+WRIST,data=data_removed)

ols_vif_tol(model_vif_2)

```

##### keep BMI, remove other variables, of which VIF above 10
```{r }
model_vif_3<-lm(BODYFAT~ AGE+ADIPOSITY+NECK+KNEE+ANKLE+BICEPS+FOREARM+WRIST,data=data_removed)

ols_vif_tol(model_vif_3)

```

## Stepwise Linear Regression
### include bmi,age,neck,knee,ankle,biceps,forearm,wrist

### Test&Train set split
```{r}
## 75% of the sample size
smp_size <- floor(0.75 * nrow(data_removed))

## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(data_removed)), size = smp_size)

train <- data_removed[train_ind, ]
test <- data_removed[-train_ind, ]

```

### Train
```{r }

library(MASS)
full_model<-lm(BODYFAT~ AGE+ADIPOSITY+NECK+KNEE+ANKLE+BICEPS+FOREARM+WRIST,data=train)

# Stepwise regression model
step_model <- stepAIC(full_model, direction = "backward", 
                      trace = FALSE)
result<-summary(step_model)
mse<-mean(result$residuals^2)
print("linear model:")
print(result)
print(paste0("MSE of train set is ",mse))

##Test
prediction<-predict(step_model,newdata=test)
mse_test=mean((test$BODYFAT-prediction)^2)
print(paste0("MSE of test set is ",mse_test))

```
### only keep variables with p-value<0.01
```{r}
simple_model_1<-lm(BODYFAT~ AGE+ADIPOSITY+WRIST,data=train)
result<-summary(simple_model_1)
mse<-mean(result$residuals^2)
print("linear model:")
print(result)
print(paste0("MSE of train set is ",mse))

##Test
prediction<-predict(simple_model_1,newdata=test)
mse_test=mean((test$BODYFAT-prediction)^2)
print(paste0("MSE of test set is ",mse_test))

```
### only keep age and bmi
```{r}
simple_model_1<-lm(BODYFAT~ AGE+ADIPOSITY,data=train)
result<-summary(simple_model_1)
mse<-mean(result$residuals^2)
print("linear model:")
print(result)
print(paste0("MSE of train set is ",mse))

##Test
prediction<-predict(simple_model_1,newdata=test)
mse_test=mean((test$BODYFAT-prediction)^2)
print(paste0("MSE of test set is ",mse_test))

```

### include height,weight,age,neck,knee,ankle,biceps,forearm,wris
```{r}
full_model_2<-lm(BODYFAT~ AGE+WEIGHT+HEIGHT+NECK+KNEE+ANKLE+BICEPS+FOREARM+WRIST,data=train)

# Stepwise regression model
step_model_2 <- stepAIC(full_model_2, direction = "backward", 
                      trace = FALSE)
result_2<-summary(step_model_2)
mse_2<-mean(result_2$residuals^2)
print("linear model:")
print(result_2)
print(paste0("MSE of train set is ",mse_2))

##Test
prediction_2<-predict(step_model_2,newdata=test)
mse_test_2=mean((test$BODYFAT-prediction_2)^2)
print(paste0("MSE of test set is ",mse_test_2))



```
### only keep variables with p-value<0.001
```{r}
simple_model_3<-lm(BODYFAT~ AGE+WEIGHT+HEIGHT+WRIST,data=train)


result<-summary(simple_model_3)
mse<-mean(result$residuals^2)
print("linear model:")
print(result)
print(paste0("MSE of train set is ",mse))

##Test
prediction<-predict(simple_model_3,newdata=test)
mse_test=mean((test$BODYFAT-prediction)^2)
print(paste0("MSE of test set is ",mse_test))

```
### Only keep age and height,weight
```{r}
simple_model_4<-lm(BODYFAT~ AGE+WEIGHT+HEIGHT,data=train)


result<-summary(simple_model_4)
mse<-mean(result$residuals^2)
print("linear model:")
print(result)
print(paste0("MSE of train set is ",mse))

##Test
prediction<-predict(simple_model_4,newdata=test)
mse_test=mean((test$BODYFAT-prediction)^2)
print(paste0("MSE of test set is ",mse_test))
```


### Final validation
```{r}
library(caret)
# Define training control
set.seed(123) 
train.control <- trainControl(method = "cv", number = 5)
# Train the model
model <- train(BODYFAT~ AGE+ADIPOSITY, data = data_removed, method = "lm",
               trControl = train.control)
# Summarize the results
print(model)

model$resample

```
